Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-05-31 appendix.tex
edited appendix

In this appendix, we provide details for the various interpolation
methods available in the code.  We assume throughout that we are
dealing with Cartesian coordinates and equally spaced cells, although
we allow for arbitrary refinement factor $r$.

@@

This interpolation algorithm is generally second-order but has
monotonicity constraints as described below.

In one dimension, we define the parent values as $Q_{-1}$, $Q_0$, and
$Q_{+1}$, where the central parent cell has a left edge at $x_0$ and
width $\Delta x$. We first linearly interpolate the parent values to
the cell edges: $Q_{-1/2} = (Q_0 + Q_{-1})/2$, and similarly for
$Q_{1/2} = (Q_0 + Q_1)/2$, and then compute a monotonic slope: $\Delta
Q_0 = \minmod{Q_{1/2} - Q_0}{Q_0 - Q_{-1/2}}$ where

@@

where $s^\prime = \min(\max(s,0), 1)$ and $\epsilon$ is a small number
to prevent numerical errors. We then use this adjustment fraction to
compute the new adjusted $\Delta Q_n$ (for $n = 1, 2, 3$) that match
the $\Delta Q_0$ constraint as closely as possible with

@@

Finally, these are converted to the grid axes with $f_x = \Delta Q_2 +
\Delta Q_3$, $f_y = \Delta Q_1 + \Delta Q_3$, $f_z = \Delta Q_1 +
\Delta Q_2$ and then we do the interpolation itself with,

@@

We also provide a variant on the above procedure, with two changes.
The first is that the slopes across the diagonals ($\Delta Q_0$, etc.)
are computed directly rather than with the minmod limiter
(e.g. $\Delta Q_0 = (Q_{1/2,1/2,1/2} - Q_{-1/2,-1/2,-1/2})/2$).  To
ensure positivity in the resulting interpolation when applied to
positive conserved quantities, the slopes are limited so that the
smallest corner value is 0.2 the cell center value.  The same
procedure as described above is then applied to turn these four slopes
into a linear interpolation.

@@

This is a completely different second-order interpolation scheme,
which is based on Cloud-In-Cell (CIC) interpolation \citep{Hockney88}.
In one dimension, we define the parent values as $Q_0$, and $Q_{+1}$,
where the left parent cell has a cell center at $x_0$ and width
$\Delta x$.  Then, the interpolated value for a subgrid cell $q_i$
with a cell left edges at $x_i = x_0 + i \Delta x^p/r$ where $i$ runs
from 0 to $r-1$ for refinement factor $r$, is given by:

@@

The extension to two and three dimensions is straightforward, with
weights in the other dimensions computed in the same way and then
multiplied to get a total of 4 and 8 weights for the 2 and 3
dimensional cases, respectively.  This scheme is monotonicity
preserving, but is not conservative.

>>>>

In this appendix, we provide details for the various interpolation
methods available in the code.  We assume throughout that we are
dealing with Cartesian coordinates and cells of equal sizes, although
we allow for arbitrary an refinement factor $r$ between cells at
different levels.

@@

This interpolation algorithm is generally second-order, but has
monotonicity constraints as described below.
In one dimension, we define the parent values as $Q_{-1}$, $Q_0$, and
$Q_{+1}$, where the central parent cell has a left edge at $x_0$ and
width $\Delta x$. We first linearly interpolate the parent values to
the cell edges: $Q_{-1/2} = (Q_0 + Q_{-1})/2$, and similarly for
$Q_{1/2} = (Q_0 + Q_1)/2$, and then compute a monotonic slope: $\Delta
Q_0 = \minmod{Q_{1/2} - Q_0}{Q_0 - Q_{-1/2}}$ where

@@

where $s^\prime = \min(\max(s,0), 1)$ and $\epsilon$ is a small number
to prevent numerical errors. We then use this adjustment fraction to
compute the new, adjusted $\Delta Q_n$ (for $n = 1, 2, 3$) that match
the $\Delta Q_0$ constraint as closely as possible with

@@

Finally, these are converted to the grid axes with $f_x = \Delta Q_2 +
\Delta Q_3$, $f_y = \Delta Q_1 + \Delta Q_3$, $f_z = \Delta Q_1 +
\Delta Q_2$.  We then do the interpolation itself with

@@

We also provide a variant on the above procedure, with two changes.
The first is that the slopes across the diagonals ($\Delta Q_0$, etc.)
are computed directly rather than with the minmod limiter
(e.g. $\Delta Q_0 = (Q_{1/2,1/2,1/2} - Q_{-1/2,-1/2,-1/2})/2$).  To
ensure positivity in the resulting interpolation when applied to
positive conserved quantities, the slopes are limited so that the
smallest corner value is 0.2 the cell center value.  The procedure
described above is then applied to turn these four slopes into a
linear interpolation.

@@

This is a completely different second-order interpolation scheme,
which is based on Cloud-In-Cell (CIC) interpolation \citep{Hockney88}.
In one dimension, we define the parent values as $Q_0$, and $Q_{+1}$,
where the left parent cell has a cell center at $x_0$ and width
$\Delta x$.  Then, the interpolated value for a subgrid cell $q_i$
with a cell left edges at $x_i = x_0 + i \Delta x^p/r$, where $i$ runs
from 0 to $r-1$ for refinement factor $r$, is given by:

@@

The extension to two and three dimensions is straightforward, with
weights in the other dimensions computed in the same way and then
multiplied to get a total of 4 and 8 weights for the 2 and 3
dimensional cases, respectively.  This scheme preserves monotonicity,
but is not conservative.
